---
output:
  knitrBootstrap::bootstrap_document:
    theme.chooser: TRUE
    highlight.chooser: TRUE
  html_document:
    toc: true
    theme: spacelab
---

Protocol template
=================


```{r bibsetup, echo=FALSE, message=FALSE, warning=FALSE}
## Load knitcitations with a clean bibliography
library('knitcitations')
cleanbib()
cite_options(hyperlink = 'to.doc', citation_format = 'text', style = 'html')

bibs <- c('knitcitations' = citation('knitcitations'),
    R = citation()
)
write.bibtex(bibs, file = 'protocolTemplate.bib')
bib <- read.bibtex('protocolTemplate.bib')
names(bib) <- names(bibs)
```

This protocol template was contributed by [L. Collado-Torres](http://bit.ly/LColladoTorres).


# Overview


TODO list:

* Check the [Markdown basics](http://rmarkdown.rstudio.com/authoring_basics.html) so you'll be familiar with the markdown syntax.
* Change the title.
* Choose your favorite HTML format. See http://rmarkdown.rstudio.com/html_document_format.html for details.
* Specify your name as the contributor and link to your webpage or GitHub profile.
* If your template includes `R` code, include the session information using `devtools::session_info()`. This is specially important if you have any packages installed from GitHub. You might want to use `options(width = 120)` before `devtools::session_info()`.
* Use [render.R](../render.R) to render your file. This will create the HTML file.


# References



Citations made with `knitcitations` `r citep(bib[['knitcitations']])`.


```{r 'bibliography', results='asis', echo=FALSE, cache=FALSE}
## Print bibliography
bibliography()
```

```{r 'cleanupBib', echo = FALSE, results = 'hide'}
## Clean up
file.remove('protocolTemplate.bib')
```

`R` information (delete this if no `R` code was used)

```{r 'rSession', echo = FALSE}
library('devtools')
options(width = 120)
session_info()
```


Date this protocol was last modified: `r Sys.time()`.
